---
title: "p8105_hw2_ht2630"
author: "Haotian Tang"
date: "`r Sys.Date()`"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

---

Problem 1

Clean data in pols_month.csv file.
```{r clean pols}

pols_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |>
  
  janitor::clean_names() |>
  
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) |>
  
  mutate(month = factor(month, labels = month.name)) |>
  
  mutate(president = case_when(
    prez_gop == 1 ~ "gop",
    prez_dem == 1 ~ "dem",
    prez_dem == 2 ~ "gop",
  )) |>
  
  select(-prez_dem, -prez_gop, -day)
```

Clean data in snp.csv file. 

```{r clean snp}

snp_df = 
  read_csv("fivethirtyeight_datasets/snp.csv") |>
  
  janitor::clean_names()|>

  separate(date, into = c("month", "day", "year"), sep = "/") |>
  
  mutate(month = as.integer(month),
         day = as.integer(day),
         year = as.integer(year)) |>
  
  mutate(year = if_else(year < 100, year + 2000, year)) |>
  
  mutate(year = if_else(year > 2023, year - 100, year)) |>
  
  mutate(month = month.name[as.integer(month)]) |>
  
  arrange(year, month) |>
  
  select(year, month, close)

```

Tidy and merge them

```{r clean unemployment}

unemployment_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |>
  
  pivot_longer(cols = -Year, names_to = "month", values_to = "unemployment_rate") |>

  rename(year = Year) |>
  
  mutate(month = match(month, month.abb),
         month = month.name[month]) |>
  
  arrange(year, month)
```

```{r merge them together}

first_merge = left_join(pols_df, snp_df, by = c("year", "month"))

merged_data = left_join(first_merge, unemployment_df, by = c("year", "month"))

```

The final data contains the specific date and pols, snp, and unemployment data. And we can also notice "NA" in the dataset, which means the data of those variables are missing.  

---

Problem 2

```{r Mr. Trash Wheel}

mr_trash_wheel_df = 
  read_excel("Trash_wheel_data.xlsx", sheet = 'Mr. Trash Wheel') |>
  
  janitor::clean_names() |>
  
  rename(dumpster_id = dumpster) |>
  
  mutate(trash_wheel_name = "Mr. Trash Wheel") |>
  
  select(1, trash_wheel_name, everything()) |>
  
  select(-x15:-x16) |>
  
  mutate(date = day(ymd(date))) |>
  
  mutate(homes_powered = (weight_tons * 500) / 30)
  
```

```{r Professor Trash Wheel}

professor_trash_wheel_df = 
  read_excel("Trash_wheel_data.xlsx", sheet = 'Professor Trash Wheel') |>
  
  janitor::clean_names() |>
  
  rename(dumpster_id = dumpster) |>
  
  mutate(year = as.character(year)) |>
  
  mutate(trash_wheel_name = "Professor Trash Wheel") |>
  
  select(1, trash_wheel_name, everything()) |>
  
  mutate(date = day(ymd(date))) |>
  
  mutate(homes_powered = (weight_tons * 500) / 30) 
  
  total_trash = professor_trash_wheel_df[nrow(professor_trash_wheel_df), "weight_tons"]
  
  total_trash = as.numeric(total_trash)
  
```

```{r Gwynnda Trash Wheel}

gwynnda_trash_wheel_df = 
  read_excel("Trash_wheel_data.xlsx", sheet = 'Gwynnda Trash Wheel') |>
  
  janitor::clean_names() |>
  
  slice(-n() + 1) |>
  
  rename(dumpster_id = dumpster) |>
  
  mutate(year = as.character(year)) |>
  
  mutate(trash_wheel_name = "Gwynnda Trash Wheel") |>
  
  select(1, trash_wheel_name, everything()) |>
  
  mutate(date = day(ymd(date))) |>
  
  mutate(homes_powered = (weight_tons * 500) / 30)

```

Combine them together

```{r Combine}

combined_df = 
  bind_rows(mr_trash_wheel_df, professor_trash_wheel_df, gwynnda_trash_wheel_df)

```

We can see there are `r nrow(mr_trash_wheel_df)` rows of data of Mr. Trash wheel, `r nrow(professor_trash_wheel_df)` rows of Professor Trash wheel, and `r nrow(gwynnda_trash_wheel_df)` rows of Gwynnda Trash Wheel. Each of the data frame contains a row which shows the total value of each column. Interestingly, there are two rows with identical dumpster id "21" in Gwynnda Trash wheel, which might be a typo. 

And there are `r nrow(combined_df)` rows of data in total. The variables of columns are dumpster, the date, the weight, the volume, types of trash, and the number of homes powered. The number of homes powered is calculated by the weights of trash multiplying 500 and divided by 30. 

The total weight of trash collected by Professor Trash Wheel is `r total_trash` tons. 

```{r}

total_cigarettes = sum(head(gwynnda_trash_wheel_df$cigarette_butts, 5))


```

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r total_cigarettes`. 

---

Problem 3 

```{r baseline}

baseline_df = 
  read_csv("data_mci/MCI_baseline.csv") 

colnames(baseline_df) = c("Study_ID", "Age_at_study_baseline", "Sex", "Years_of_Education", "Apoe4_carrier", "Age_at_onset")

baseline_df = slice(baseline_df, 2:n()) |>

  mutate(Sex = case_when(
    Sex == 0 ~ "male",
    Sex == 1 ~ "female",
    TRUE ~ as.character(Sex)
  )) |> 
  
    mutate(Apoe4_carrier = case_when(
    Apoe4_carrier == 0 ~ "No",
    Apoe4_carrier == 1 ~ "Yes",
    TRUE ~ as.character(Apoe4_carrier)
  )) 
  
  total_females = sum(baseline_df$Sex == "female")
  
  female_apoe4_carriers = sum(baseline_df$Sex == "female" & baseline_df$`Apoe4_carrier` == "Yes")
  
  proportion = female_apoe4_carriers / total_females

  baseline_df |>
    
  filter(`Age_at_onset` != ".")
  
  age = as.numeric(baseline_df$Age_at_study_baseline) 
    
  mean(age)
  
  baseline_candidates = baseline_df[baseline_df$`Age_at_onset` != ".", ]
  
  mci_candidates = sum(baseline_df$Age_at_onset != ".")

```


First, the column names need to be adjusted since the first row is the notes of column variables. 

And then we need to remove the first row. 

There are `r nrow(baseline_df)` rows and `r ncol(baseline_df)` columns. 

Variables are `r colnames(baseline_df)`. 

And we remove patients who don't have MCI at baseline. 

There are `r nrow(baseline_df)` participants but only `r mci_candidates` of them develop MCI. The average age is `r mean(age)`. The proportion of women in the study who are Apoe4 carrier is `r proportion`. 

```{r amyloid}

amyloid_df = 
  read_csv("data_mci/mci_amyloid.csv") 

colnames(amyloid_df) = c("Study_ID", "Baseline", "Year_2", "Year_4", "Year_6", "Year_8")

amyloid_df = slice(amyloid_df, 2:n()) |> 
  
  mutate(across(c("Baseline", "Year_2", "Year_4", "Year_6", "Year_8"), ~ replace(., . == "Na", NA))) |>
  
  pivot_longer(
    Baseline:Year_8,
    names_to = "year",
    values_to = "value"
  )
  

```

Similarly, the column names need to be changed and the first row is removed. 

There are `r nrow(amyloid_df)` rows and `r ncol(amyloid_df)` columns. 

Variables are `r colnames(amyloid_df)`.

And we remove patients who have NA at any year. 

Since the rows are not equal, there are patients who are only in one dataset. 

```{r Merge them}

combined_AD_df = inner_join(baseline_candidates, amyloid_df, by = "Study_ID")

```
There are `r nrow(combined_AD_df)` rows and `r ncol(combined_AD_df)` columns.

Variables are `r colnames(combined_AD_df)`.

So only `r nrow(combined_AD_df)` patients have complete data about baseline and amyloid. 

The merged dataset is very easy and straightforward to perform data analysis. 

```{r}
write.csv(combined_AD_df, file = "Merged_data.csv", row.names = FALSE)
```



